exp1_lik_by_trial

### 2020-09-04 comp_model_1019.Rmd
fixed Q learning bugs

#Fit a bayesian model to observed human play. Prior is vector of possible level-k opponent strategies.
```{r}
# Function to produce best reponse to ai_action 
best_response <- function(game,ai_action){
  if (game == "rps"){
    if (ai_action == "R") {return("P")}
    else if (ai_action == "P") {return("S")}
    else if (ai_action == "S") {return("R")}
  } else if (game == "fwg"){ 
    if (ai_action == "F") {return("W")}
    else if (ai_action == "W") {return("G")}
    else if (ai_action == "G") {return("F")}
  } else if (game == "numbers") {
    if (ai_action == "1") {return("2")}
    else if (ai_action == "2") {return("3")}
    else if (ai_action == "3") {return("4")}
    else if (ai_action == "4") {return("5")}
    else if (ai_action == "5") {return("1")}
  }
}

exp1_ToM <- function(par,data,opp_strategy_vec,return_value,opp_mod_transfer) {

  # dat = data subset for one participant
  # opp_strategy_vec = vector of possible opponent strtegies, model assumes humans restrict opp strategy space to vector   c("level0","level1","level2")
  # return = -2logLik 
  # or "likelihood_by_trial" vector  (for later plotting, etc)
  # opp_mod_transfer is a boolean. True if distribution of opponent strategies is kept across games. False otherwise.

  # theta = parameter, probability computer will play its best response to what it thinks human is. Truth = 90%. in [0,1].
  # eps: parameter controlling noise in human choice. Probability human will DEVIATE from its best response. In [0,1].
  theta <- par[1]
  eps <- par[2]

  num_strat = length(opp_strategy_vec)

  # Load tables predicting oppponent move
  rps_predict_opp <- read.csv("rps_predict_opp.csv")
  fwg_predict_opp <- read.csv("fwg_predict_opp.csv")
  numbers_predict_opp <- read.csv("numbers_predict_opp.csv")

  #lik_opp is a vector that holds and update probability distribution of opp strategy given actions.
  lik_opp <- matrix(0.0,num_strat, nrow(data))
  # Initiate prior vector as uniform
  opp_prior_vec <-rep(1/num_strat, num_strat)

  lik_hum <- matrix(0.0,nrow(data))
  # br_hum is vector that stores best responses of human to actions taken by each level-k ai opponent at the round.
  br_hum <- rep(NA,num_strat)

  for(t in 1:nrow(data)) {
    t_game <- data[t,"game"]
    pred_file_opp <- switch(as.character(t_game),rps = rps_predict_opp,fwg = fwg_predict_opp, numbers = numbers_predict_opp)

    if(t_game == "numbers") nopts <- 5 else nopts <- 3
    if(data[t,"round"] == 1) {
      # first round is uniform prediction
      pred_opp <- NA
      lik_hum[t] <- 1/nopts
      ## KEY CODE FOR LEARNING TRANSFER !! (if no OM transfer, reset prior to uniform)
      if (!(opp_mod_transfer)) {
        opp_prior_vec <- rep(1/num_strat, num_strat)
        lik_opp[,t] <- opp_prior_vec
        }

    # Rounds after first
    } else {
      # get prediction of opponent action from CSV files
      for (strategy in opp_strategy_vec) {
        k = match(strategy,opp_strategy_vec)
        pred_opp <- as.character(filter(pred_file_opp, pred_file_opp$human_previous == as.character(data[t-1,"h_action"]) & pred_file_opp$computer_previous == as.character(data[t-1,"a_action"]))[[strategy]])

        # Given opponent action predicted, what would be human action best response for each opp strat
        br_hum[[k]] <- best_response(t_game,as.character(pred_opp))

        # Multiply prior by likelihood of observation to get posterior. Done here ot take advantage of For loop. 
        if(as.character(data[t,"a_action"]) == pred_opp) {
          lik_opp[k,t] <- (theta + (1-theta)/nopts)*opp_prior_vec[k]
        } else {
          lik_opp[k,t] <- ((1-theta)/nopts)*opp_prior_vec[k]
        }
      }
      # Get which opponent strategy human current actn maps to (if index = 0 then human current action is br to level-0 comp_strat...)
      indices <- which(br_hum %in% as.character(data[t,"h_action"]))

      # THEN likelihood of current human action is just prior on opponent vec (haven't updated priors with curr opp act yet)
      # if action is not predicted by any level-k OM, assume human chooses randomly with prob eps
      if(length(indices) == 0){
        lik_hum[t] <- eps/nopts
      # if same br action for multiple possible ai opponents, add probabilities 
      } else {
        lik_hum[t] <- sum(opp_prior_vec[indices])*(1-eps) + eps/nopts
      }
      # Standardising the probabilities and updating prior
      lik_opp[,t] <- lik_opp[,t]/sum(lik_opp[,t])
      opp_prior_vec <- lik_opp[,t]
    }
    # for debugging uncomment the line below
    #cat(as.character(pred_opp),as.character(data[t,"a_action"]),lik_opp[,t],"\n")
  }
  if(return_value == "-2loglik") {
    ret <- -2*sum(log(lik_hum))
    if(is.infinite(ret) || is.nan(ret)) {
      return(1e+300)
    } else {
      return(ret)
    }
  }
  if(return_value == "likelihood_by_trial") return(lik_hum)
}

# data = subset(dat,human_id == "38VxtUSv_h6RR5-tAAA2")
# exp1_ToM(c(0.9,0.99),data,c("level0","level1","level2"),"-2loglik",FALSE)
```

```{r}
# rps_predict_opp <- read.csv("rps_predict_opp.csv")
# red_opp <- filter(rps_predict_opp, rps_predict_opp$human_previous == "R" & rps_predict_opp$computer_previous == "S")
# 
# as.character(red_opp[["level0"]])
```

# Fitting a bayesian model WITH BETWEEN GAME TRANSFER of opp strat distribution between games 
```{r, cache=TRUE}
exp1_ToM_TR <- list()
for(id in unique(dat$human_id)) {
  exp1_ToM_TR[[id]] <- list()
  tdat <- subset(dat,human_id == id)

  # exp1_ToM_TR[[id]] <- DEoptim(fn=exp1_ToM, lower = c(0,0), upper = c(1,1), data=tdat, opp_strategy_vec = c("level0","level1","level2") ,return_value = "-2loglik", opp_mod_transfer = TRUE, control=list(trace = FALSE, parallelType=1,parVar = c("best_response")))

  exp1_ToM_TR[[id]] <- optim(c(0.1,0.1),fn=exp1_ToM,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = TRUE,lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")
}
save(exp1_ToM_TR,file="exp1_ToM_TR.RData")

@@

# Fitting a bayesian model with NO BETWEEN GAME TRANSFER of opp strat distribution between games 
```{r, cache =TRUE}
exp1_ToM_NT <- list()
for(id in unique(dat$human_id)) {
  exp1_ToM_NT[[id]] <- list()
  tdat <- subset(dat,human_id == id)

  # exp1_ToM_NT[[id]] <- DEoptim(fn=exp1_ToM, lower = c(0.0,0.0), upper = c(1.0,1.0), data=tdat, opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = FALSE, control=list(trace = FALSE, parallelType=0,parVar = c("best_response")))

  exp1_ToM_NT[[id]] <- optim(c(0.1,0.1),fn=exp1_ToM,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = FALSE ,lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")
}

save(exp1_ToM_NT,file="exp1_ToM_NT.RData")

# Summarise results in dataframe for both transfer and no transfer models 
```{r}
Nbayes_par_results <- data.frame()
for(id in unique(dat$human_id)) {
Nbayes_par_results <- rbind(Nbayes_par_results,
                       data.frame(
                         "id" = id,
                         "condition" = dat[dat$human_id==id,"condition"][1],
                         # BICs 
                         "BIC_transfer" = exp1_ToM_TR[[id]]$value + 2*log(150),
                         "BIC_no_transfer" = exp1_ToM_NT[[id]]$value+ 2*log(150),
                         # Theta is the parameter governing AI stochasticity. Truth is 0.9
                         "theta_transfer" = exp1_ToM_TR[[id]]$par[1],
                         "theta_no_transfer" = exp1_ToM_NT[[id]]$par[1],
                         # Epsilone is parameter showing cases where human deviates from predictions about opponent (lower better)
                         "eps_transfer" = exp1_ToM_TR[[id]]$par[2],
                         "eps_no_transfer" = exp1_ToM_NT[[id]]$par[2]))
}

#Plotting the theta/eps scatter for all participants...
plot(Nbayes_par_results$theta_transfer,Nbayes_par_results$eps_transfer, xlab="Theta", ylab= "Epsilone",col = ifelse(as.character(Nbayes_par_results$condition) == "Level2",'red','green'))
# 
write.csv(Nbayes_par_results,file="Nbayes_par_results.csv",row.names = FALSE)
```

@@

## Q-Learning with state space consisting of last round play
```{r}

Q_learn_states <- function(par,data,return_value,gamma){
  # Par[1] -> beta= inverse temperature parameter in softmax choice function
  # Par[2] -> lambda = learning rate (one per game?)
  # data : participant and ai choice data.
  # return_value=c("-2loglik","likelihood_by_trial")
  # Gamma is discount factor for future rewards 
  # Returns Q-values per trial and predicts choice using softmax
  beta <- par[1]
  alpha <- par[2]

@@

    if(data[t,"round"] == 1) {
      # first round is uniform prediction
      Q_vals <- switch(as.character(t_game),rps=Q_vals_RPS,fwg = Q_vals_FWG, numbers = Q_vals_NUM)
      state_vec <- switch(as.character(t_game),rps=states_RPS,fwg = states_FWG, numbers = states_NUM)
      lik_hum[t] <- 1/nopts
      # Randomly select prev_state and actions for first round
      curr_state <- sample(state_vec, size = 1)
      h_act <- sample(colnames(Q_vals), size =1)
      ai_act <- sample(colnames(Q_vals), size =1)
      reward <- 0

    # Q_learning: update rule (time  = t-1)
    Q_vals[curr_state, h_act] <- Q_vals[curr_state, h_act] + alpha*( reward + gamma*max(Q_vals[new_state,]) - Q_vals[curr_state, h_act])

    # Update state
    curr_state <- new_state 
    #}
  }
  if(return_value == "-2loglik") {
    ret <- -2*sum(log(lik_hum))
    if(is.infinite(ret) || is.nan(ret)) {
      return(1e+300)
    } else {
      return(ret)
    }
  }
  if(return_value == "likelihood_by_trial") return(lik_hum)

@@

@@

write.csv(All_results,file="All_results.csv",row.names = FALSE)

>>>>

## Fit a bayesian model to observed human play. Prior is vector of possible level-k opponent strategies.
# # Function to produce best reponse to ai_action 
# best_response <- function(game,ai_action){
#   if (game == "rps"){
#     if (ai_action == "R") {return("P")}
#     else if (ai_action == "P") {return("S")}
#     else if (ai_action == "S") {return("R")}
#   } else if (game == "fwg"){ 
#     if (ai_action == "F") {return("W")}
#     else if (ai_action == "W") {return("G")}
#     else if (ai_action == "G") {return("F")}
#   } else if (game == "numbers") {
#     if (ai_action == "1") {return("2")}
#     else if (ai_action == "2") {return("3")}
#     else if (ai_action == "3") {return("4")}
#     else if (ai_action == "4") {return("5")}
#     else if (ai_action == "5") {return("1")}
#   }
# }
# 
# exp1_ToM <- function(par,data,opp_strategy_vec,return_value,opp_mod_transfer) {
#   
#   # dat = data subset for one participant
#   # opp_strategy_vec = vector of possible opponent strtegies, model assumes humans restrict opp strategy space to vector   c("level0","level1","level2")
#   # return = -2logLik 
#   # or "likelihood_by_trial" vector  (for later plotting, etc)
#   # opp_mod_transfer is a boolean. True if distribution of opponent strategies is kept across games. False otherwise. 
#   
#   # theta = parameter, probability computer will play its best response to what it thinks human is. Truth = 90%. in [0,1].
#   # eps: parameter controlling noise in human choice. Probability human will DEVIATE from its best response. In [0,1].
#   theta <- par[1]
#   eps <- par[2]
#   
#   num_strat = length(opp_strategy_vec) 
#   
#   # Load tables predicting oppponent move
#   rps_predict_opp <- read.csv("rps_predict_opp.csv")
#   fwg_predict_opp <- read.csv("fwg_predict_opp.csv")
#   numbers_predict_opp <- read.csv("numbers_predict_opp.csv")
# 
#   #lik_opp is a vector that holds and update probability distribution of opp strategy given actions.
#   lik_opp <- matrix(0.0,num_strat, nrow(data))
#   # Initiate prior vector as uniform
#   opp_prior_vec <-rep(1/num_strat, num_strat)
#   
#   lik_hum <- matrix(0.0,nrow(data))
#   # br_hum is vector that stores best responses of human to actions taken by each level-k ai opponent at the round.
#   br_hum <- rep(NA,num_strat)
# 
#   
#   for(t in 1:nrow(data)) {
#     t_game <- data[t,"game"]
#     pred_file_opp <- switch(as.character(t_game),rps = rps_predict_opp,fwg = fwg_predict_opp, numbers = numbers_predict_opp)
#     
#     if(t_game == "numbers") nopts <- 5 else nopts <- 3
#     if(data[t,"round"] == 1) {
#       # first round is uniform prediction
#       pred_opp <- NA
#       lik_hum[t] <- 1/nopts
#       ## KEY CODE FOR LEARNING TRANSFER !! (if no OM transfer, reset prior to uniform)
#       if (!(opp_mod_transfer)) {
#         opp_prior_vec <- rep(1/num_strat, num_strat)
#         lik_opp[,t] <- opp_prior_vec
#         }
#       
#     # Rounds after first
#     } else {
#       # get prediction of opponent action from CSV files
#       for (strategy in opp_strategy_vec) {
#         k = match(strategy,opp_strategy_vec)
#         pred_opp <- as.character(filter(pred_file_opp, pred_file_opp$human_previous == as.character(data[t-1,"h_action"]) & pred_file_opp$computer_previous == as.character(data[t-1,"a_action"]))[[strategy]])
#         
#         # Given opponent action predicted, what would be human action best response for each opp strat
#         br_hum[[k]] <- best_response(t_game,as.character(pred_opp))
#         
#         # Multiply prior by likelihood of observation to get posterior. Done here ot take advantage of For loop. 
#         if(as.character(data[t,"a_action"]) == pred_opp) {
#           lik_opp[k,t] <- (theta + (1-theta)/nopts)*opp_prior_vec[k]
#         } else {
#           lik_opp[k,t] <- ((1-theta)/nopts)*opp_prior_vec[k]
#         }
#       }
#       # Get which opponent strategy human current actn maps to (if index = 0 then human current action is br to level-0 comp_strat...)
#       indices <- which(br_hum %in% as.character(data[t,"h_action"]))
# 
#       # THEN likelihood of current human action is just prior on opponent vec (haven't updated priors with curr opp act yet)
#       # if action is not predicted by any level-k OM, assume human chooses randomly with prob eps
#       if(length(indices) == 0){
#         lik_hum[t] <- eps/nopts
#       # if same br action for multiple possible ai opponents, add probabilities 
#       } else {
#         lik_hum[t] <- sum(opp_prior_vec[indices])*(1-eps) + eps/nopts
#       }
#       # Standardising the probabilities and updating prior
#       lik_opp[,t] <- lik_opp[,t]/sum(lik_opp[,t])
#       opp_prior_vec <- lik_opp[,t]
#     }
#     # for debugging uncomment the line below
#     #cat(as.character(pred_opp),as.character(data[t,"a_action"]),lik_opp[,t],"\n")
#   }
#   if(return_value == "-2loglik") {
#     ret <- -2*sum(log(lik_hum))
#     if(is.infinite(ret) || is.nan(ret)) {
#       return(1e+300)
#     } else {
#       return(ret)
#     }
#   }
#   if(return_value == "likelihood_by_trial") return(lik_hum)
# }
# 
# # data = subset(dat,human_id == "38VxtUSv_h6RR5-tAAA2")
# # exp1_ToM(c(0.9,0.99),data,c("level0","level1","level2"),"-2loglik",FALSE)
# ```
# 
# ```{r}
# # rps_predict_opp <- read.csv("rps_predict_opp.csv")
# # red_opp <- filter(rps_predict_opp, rps_predict_opp$human_previous == "R" & rps_predict_opp$computer_previous == "S")
# # 
# # as.character(red_opp[["level0"]])
# ```
# 
# 
# 
# # Fitting a bayesian model WITH BETWEEN GAME TRANSFER of opp strat distribution between games 
# ```{r, cache=TRUE}
# exp1_ToM_TR <- list()
# for(id in unique(dat$human_id)) {
#   exp1_ToM_TR[[id]] <- list()
#   tdat <- subset(dat,human_id == id)
#   
#   # exp1_ToM_TR[[id]] <- DEoptim(fn=exp1_ToM, lower = c(0,0), upper = c(1,1), data=tdat, opp_strategy_vec = c("level0","level1","level2") ,return_value = "-2loglik", opp_mod_transfer = TRUE, control=list(trace = FALSE, parallelType=1,parVar = c("best_response")))
# 
#   exp1_ToM_TR[[id]] <- optim(c(0.1,0.1),fn=exp1_ToM,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = TRUE,lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")
# }
# save(exp1_ToM_TR,file="exp1_ToM_TR.RData")
# 
# ```
# 
# 
# # Fitting a bayesian model with NO BETWEEN GAME TRANSFER of opp strat distribution between games 
# ```{r, cache =TRUE}
# exp1_ToM_NT <- list()
# for(id in unique(dat$human_id)) {
#   exp1_ToM_NT[[id]] <- list()
#   tdat <- subset(dat,human_id == id)
#   
#   # exp1_ToM_NT[[id]] <- DEoptim(fn=exp1_ToM, lower = c(0.0,0.0), upper = c(1.0,1.0), data=tdat, opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = FALSE, control=list(trace = FALSE, parallelType=0,parVar = c("best_response")))
# 
#   exp1_ToM_NT[[id]] <- optim(c(0.1,0.1),fn=exp1_ToM,gr = NULL, data=tdat,opp_strategy_vec = c("level0","level1","level2"),return_value = "-2loglik", opp_mod_transfer = FALSE ,lower = c(0.01,0.01), upper = c(0.99,0.99), method="L-BFGS-B")
# }
# 
# save(exp1_ToM_NT,file="exp1_ToM_NT.RData")
# 
# ```
# 
# 
# # Summarise results in dataframe for both transfer and no transfer models 
# ```{r}
# Nbayes_par_results <- data.frame()
# for(id in unique(dat$human_id)) {
# Nbayes_par_results <- rbind(Nbayes_par_results,
#                        data.frame(
#                          "id" = id,
#                          "condition" = dat[dat$human_id==id,"condition"][1],
#                          # BICs 
#                          "BIC_transfer" = exp1_ToM_TR[[id]]$value + 2*log(150),
#                          "BIC_no_transfer" = exp1_ToM_NT[[id]]$value+ 2*log(150),
#                          # Theta is the parameter governing AI stochasticity. Truth is 0.9
#                          "theta_transfer" = exp1_ToM_TR[[id]]$par[1],
#                          "theta_no_transfer" = exp1_ToM_NT[[id]]$par[1],
#                          # Epsilone is parameter showing cases where human deviates from predictions about opponent (lower better)
#                          "eps_transfer" = exp1_ToM_TR[[id]]$par[2],
#                          "eps_no_transfer" = exp1_ToM_NT[[id]]$par[2]))
# }
# 
# #Plotting the theta/eps scatter for all participants...
# plot(Nbayes_par_results$theta_transfer,Nbayes_par_results$eps_transfer, xlab="Theta", ylab= "Epsilone",col = ifelse(as.character(Nbayes_par_results$condition) == "Level2",'red','green'))
# # 
# write.csv(Nbayes_par_results,file="Nbayes_par_results.csv",row.names = FALSE)

@@

@@

## First attempt : Q-Learning with state space consisting of last round play (deprecated) 
```{r}

# Q_learn_states <- function(par,data,return_value,gamma){
#   # Par[1] -> beta= inverse temperature parameter in softmax choice function
#   # Par[2] -> lambda = learning rate (one per game?)
#   # data : participant and ai choice data.
#   # return_value=c("-2loglik","likelihood_by_trial")
#   # Gamma is discount factor for future rewards 
#   # Returns Q-values per trial and predicts choice using softmax
#   beta <- par[1]
#   alpha <- par[2]
# 
#   #gamma <- 0.9 
# 
#   #Define matrix of state spaces for each game 
#   G1 <- expand.grid(c("R", "P", "S"),c("R", "P", "S"))
#   states_RPS <- paste0(G1$Var1,G1$Var2)
# 
#   G2 <- expand.grid(c("F", "W", "G"), c("F", "W", "G"))
#   states_FWG <- paste0(G2$Var1,G2$Var2)
# 
#   G3 <- expand.grid(c("1", "2", "3", "4", "5"), c("1", "2", "3", "4", "5"))
#   states_NUM <- paste0(G3$Var1,G3$Var2)
#   
#   Q_vals_RPS = matrix(-0.5,9,3)
#   dimnames(Q_vals_RPS) = list(states_RPS, c("R", "P", "S"))
# 
#   Q_vals_FWG = matrix(-0.5,9,3)
#   dimnames(Q_vals_FWG) = list(states_FWG, c("F", "W", "G"))
#   
#   Q_vals_NUM = matrix(-0.5,25,5)
#   dimnames(Q_vals_NUM) = list(states_NUM, c("1", "2", "3", "4", "5"))
#  
#   lik_hum <- matrix(0.0,nrow(data))
# 
#   for(t in 1:nrow(data)) {
#     t_game <- data[t,"game"]
#     if(t_game == "numbers") nopts <- 5 else nopts <- 3
#     
#     if(data[t,"round"] == 1) {
#       # first round is uniform prediction
#       Q_vals <- switch(as.character(t_game),rps=Q_vals_RPS,fwg = Q_vals_FWG, numbers = Q_vals_NUM)
#       state_vec <- switch(as.character(t_game),rps=states_RPS,fwg = states_FWG, numbers = states_NUM)
#       lik_hum[t] <- 1/nopts
#       # Randomly select prev_state and actions for first round
#       curr_state <- sample(state_vec, size = 1)
#       h_act <- sample(colnames(Q_vals), size =1)
#       ai_act <- sample(colnames(Q_vals), size =1)
#       reward <- 0 
#       
#     } else {
#       # Get past human action and associated reward 
#       h_act <- as.character(data[t-1,"h_action"])
#       ai_act <- as.character(data[t-1,"a_action"])
#       reward <- as.numeric(data[t-1,"score"])
#     }
#     
#     
#     # cat(curr_state, " This is the previous state", "\n")
#     new_state <- paste0(h_act,ai_act)
#     # cat(new_state,"This is new state","\n")
#     
#     # Q_learning: update rule (time  = t-1)
#     Q_vals[curr_state, h_act] <- Q_vals[curr_state, h_act] + alpha*( reward + gamma*max(Q_vals[new_state,]) - Q_vals[curr_state, h_act])
#     
#     # Assume human chooses action probabilistically using softmax on Q values
#     probs <- exp(Q_vals[new_state,]/beta)/sum(exp(Q_vals[new_state,]/beta))
#     #if (data[t,"round"] == 50) { cat(Q_vals,"\n") }
#     
#     # Get actual human action and compute likelihood
#     h_act <- as.character(data[t,"h_action"])
#     act_index <- match(h_act, colnames(Q_vals))
#     lik_hum[t] <- probs[[act_index]]
#       
#     # Update state
#     curr_state <- new_state 
#     #}
#   }
#   if(return_value == "-2loglik") {
#     ret <- -2*sum(log(lik_hum))
#     if(is.infinite(ret) || is.nan(ret)) {
#       return(1e+300)
#     } else {
#       return(ret)
#     }
#   }
#   if(return_value == "likelihood_by_trial") return(lik_hum)
#   
# }

#Second attempt at QL algorithm with states: No guessing of initial states
```{r}
Q_learn_states <- function(par,data,return_value,gamma){
  # Par[1] -> beta= inverse temperature parameter in softmax choice function
  # Par[2] -> lambda = learning rate (one per game?)
  # data : participant and ai choice data.
  # return_value=c("-2loglik","likelihood_by_trial")
  # Gamma is discount factor for future rewards 
  # Returns Q-values per trial and predicts choice using softmax
  beta <- par[1]
  alpha <- par[2]

@@

  for(t in 1:nrow(data)) {

    t_game <- data[t,"game"]
    if(t_game == "numbers") nopts <- 5 else nopts <- 3

    if(data[t,"round"] == 1) {
      # first round is uniform prediction
      Q_vals <- switch(as.character(t_game),rps=Q_vals_RPS,fwg = Q_vals_FWG, numbers = Q_vals_NUM)
      state_vec <- switch(as.character(t_game),rps=states_RPS,fwg = states_FWG, numbers = states_NUM)
      lik_hum[t] <- 1/nopts

    } else {
      # Get past human action and associated reward 
      h_act_prev <- as.character(data[t-1,"h_action"])
      ai_act_prev <- as.character(data[t-1,"a_action"])
      curr_state <- paste0(h_act_prev,ai_act_prev)

      h_act <- as.character(data[t,"h_action"])
      ai_act <- as.character(data[t,"a_action"])
      reward <- as.numeric(data[t,"score"])

      new_state <- paste0(h_act,ai_act)

      # Assume human chooses action probabilistically using softmax on Q values
      probs <- exp(Q_vals[curr_state,]/beta)/sum(exp(Q_vals[curr_state,]/beta))
      # Get actual human next action and compute likelihood
      act_index <- match(h_act, colnames(Q_vals))

      lik_hum[t] <- probs[[act_index]]

       # Q_learning: update rule (time = t)
      Q_vals[curr_state, h_act] <- Q_vals[curr_state, h_act] + alpha*( reward + gamma*max(Q_vals[new_state,]) - Q_vals[curr_state, h_act])

      # Update state
      curr_state <- new_state

@@

  }
  if(return_value == "-2loglik") {
    ret <- -2*sum(log(lik_hum))
    if(is.infinite(ret) || is.nan(ret)) {
      return(1e+300)
    } else {
      return(ret)
    }
  }
  if(return_value == "likelihood_by_trial") return(lik_hum)

```{r}
# data34 <- subset(dat,human_id == "QSuzBXpbyRc370HsAACW")
# Q_learn_states(c(5, 0.5),data34 ,return_value = "likelihood_by_trial",gamma = 0 )

@@

write.csv(All_results,file="exp1_all_results.csv",row.names = FALSE)

### 2020-09-04 exp1_lik_by_trial.Rmd
added lik by trial exp 2
fixed Q learning bugs

dat <- read.csv("exp1_data.csv")
#####
# dat <- subset(dat, human_id == "38VxtUSv_h6RR5-tAAA2")
####

optim_results <- read.csv("All_results.csv")

@@

# dat <- as_tibble(dat) %>% group_by(human_id,game)
dat <- dat %>%
  mutate(ai_action_prev = lag(ai_action,1), human_action_prev = lag(human_action,1))

@@

```{r}
# All_results <- data.frame()
# for(id in unique(dat$human_id)) {
#   All_results <- rbind(All_results,
#                        data.frame(
#                          "ID" = id,
#                          "condition" = dat[dat$human_id==id,"condition"][1],
#                          "Random_BIC" = -2*(100*log(1/3) + 50*log(1/5)),

lik_by_trial <- data.frame()

@@

  # QLS likelihoods 
  pars_QLS <- unlist(filter(optim_results, ID == id)%>%select(QL_states_alpha,QL_states_Beta))
  tdat$QLS_lik <- Q_learn_states(pars_QLS,tdat,"likelihood_by_trial",gamma = 0)

  # Bayes model no transfer likelihoods
  theta_NT <- unlist(filter(optim_results, ID == id)%>%select(theta_no_transfer))
  tdat$Bayes_NT_lik <- exp1_Bayes_model_LL(theta_NT ,data = tdat, generalize = "no", "likelihood_by_trial")

  # Bayes model with Transfer likelihood
  theta_Tr <- unlist(filter(optim_results, ID == id)%>%select(theta_transfer))
  tdat$Bayes_Tr_lik <- exp1_Bayes_model_LL(theta_Tr ,data = tdat, generalize = "game", "likelihood_by_trial")

  lik_by_trial <- rbind(lik_by_trial,tdat)
}
```

```{r}
library(ggplot2)

mean_lik <- lik_by_trial %>%
  select(human_id,condition,game, round,QLS_lik,Bayes_NT_lik,Bayes_Tr_lik) %>%
  group_by(condition,game,round) %>%
  summarise(mean_QLS_lik = mean(QLS_lik, na.rm = TRUE),
            mean_NT_lik = mean(Bayes_NT_lik, na.rm = TRUE),
            mean_Tr_lik = mean(Bayes_Tr_lik, na.rm = TRUE))

>>>>

dat <- read.csv("data20180719.csv")
#####
# dat <- subset(dat, human_id == "38VxtUSv_h6RR5-tAAA2")
####

optim_results <- read.csv("exp1_all_results.csv")

@@

# transform 'winner' variable in numeric score
dat$score <- recode(dat$winner, human = 1, tie = 0, ai = -1)
# create a new variable 'block' with round 1...25 = block 1 and round 26...50 as block 2
dat$block <- as.numeric(cut(dat$round,2))

# recode actions to make them equal to the codes in these files
dat$h_action <- recode(dat$human_action,"rock" = "R", "paper" = "P", "scissors" = "S", "fire" = "F", "water" = "W", "grass" = "G", "one" = "1", "two" = "2", "three" = "3", "four" = "4", "five" = "5")
dat$a_action <- recode(dat$ai_action,"rock" = "R", "paper" = "P", "scissors" = "S", "fire" = "F", "water" = "W", "grass" = "G", "one" = "1", "two" = "2", "three" = "3", "four" = "4", "five" = "5")

# Prepping data for Bayesian code
dat <- as_tibble(dat) %>% group_by(human_id,game)
dat <- dat %>%
  mutate(ai_action_prev = lag(ai_action,1), human_action_prev = lag(human_action,1))

@@

dat <- as.data.frame(dat)                       
exp1_lik_by_trial <- data.frame()

@@

  # QLS likelihoods 
  pars_QLS <- unlist(filter(optim_results, ID == id)%>%select(QL_states_Beta,QL_states_alpha))
  QLS_lik <- Q_learn_states(pars_QLS ,tdat,"likelihood_by_trial",gamma = 0)

  # Bayes model no transfer likelihoods
  theta_NT <- unlist(filter(optim_results, ID == id)%>%select(theta_no_transfer))
  Bayes_NT_lik <- exp1_Bayes_model_LL(theta_NT ,tdat, generalize = "no", "likelihood_by_trial")

  # Bayes model with Transfer likelihood
  theta_Tr <- unlist(filter(optim_results, ID == id)%>%select(theta_transfer))
  Bayes_Tr_lik <- exp1_Bayes_model_LL(theta_Tr ,data = tdat, generalize = "game", "likelihood_by_trial")

  exp1_lik_by_trial <- rbind(exp1_lik_by_trial,
                        data.frame(
                          "human_id" = id,
                          "condition" = dat$condition,
                          "game" = dat$game,
                          "round" = dat$round,
                          "QLS_lik" = QLS_lik,
                          "Bayes_NT_lik" = Bayes_NT_lik,
                          "Bayes_Tr_lik" = Bayes_Tr_lik

                        ))
}

which(exp1_lik_by_trial$QLS_lik == 1.0 )

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)

mean_lik <- exp1_lik_by_trial %>%
  select(human_id,condition,game, round,QLS_lik,Bayes_NT_lik,Bayes_Tr_lik) %>%
  group_by(condition,game,round) %>%
  summarise(mean_QLS_lik = mean(QLS_lik, na.rm = TRUE),
            mean_NT_lik = mean(Bayes_NT_lik, na.rm = TRUE),
            mean_Tr_lik = mean(Bayes_Tr_lik, na.rm = TRUE))
